%!TEX root = ../report.tex
\documentclass[../report.tex]{subfiles}

\begin{document}
    \begin{abstract}
        Forests present a highly heterogeneous environment composed of terrain, vegetation, tree trunks, tree stumps, and other natural elements. The irregularity of Earth’s topography, combined with dense vegetation and occlusions, poses a significant challenge to distinguish terrain from other environmental components, particularly when using traditional machine learning approaches. A major bottleneck in this task lies in the feature extraction process from noisy and partially labeled 3D point-cloud datasets, which often lack semantic consistency and spatial clarity. This report proposes a robust and adaptable pipeline for semantic segmentation of 3D point clouds, using 2D annotations to guide learning. Our method integrates state-of-the-art feature extraction techniques and clustering algorithms to isolate tree stumps and terrain as independent semantic clusters. The proposed approach achieves a classification accuracy of 59.73\% for terrain and 86.77\% for segmentation in a forest UAV dataset of benchmark containing more than 2.5 million points. The results of this work are highly applicable to forest monitoring based on UAVs, with potential extensions to vegetation indexing, biomass estimation, and land use analysis in both forested and urban settings. The pipeline not only addresses the limitations of existing methods, but also sets a foundation for scalable, real-time environmental perception in ecological research and smart forestry initiatives.
    \end{abstract}
\end{document}
